%!TEX root = ../thesis.tex
\chapter{Notation Table}

%\begin{table}[t]
%	\centering
%	\begin{tabular}{ ll }
\begin{longtable}{ll}
		\toprule
		\textbf{Symbol} & \textbf{Description} \\
		\midrule
		$a_{fix}$ & Fixed selling price. \\
		$\delta$ & Discount factor for future profits. \\
		$l$ & Holding costs per item and period. \\
		$t$ & Identifier variable for a time period. \\
		$N_t$ & Random inventory level at the start of period $t$. \\
		$b_t$ & Number of items ordered at start of period $t$. \\
		$B$ & Set of admissible order quantities. \\
		$c_{fix}$ & Fixed order costs. \\
		$c_{var}$ & Variable order costs. \\
		$C(b)$ & Total order costs for ordering $b$ items. \\
		$P(i)$ & Probability to sell i items in one time period. \\
		% P(i,a)
		$X_t$ & Random number of sold items in period $(t-1, t)$. \\
		$G_t$ & Random accumulated discounted profit from period $t$. \\
		$n$ & Inventory level. \\
		$V_t(n)$ & \makecell[l]{Value function: best expected discounted profit from \\ period $t$ with a start inventory of $n$ items.} \\
		$T$ & Number of periods of the limited time horizon. \\
		$N_{max}$ & Maximum inventory capacity. \\
		$b$ & Order quantity. \\
		$u$ & Start value for value function: $V_T(n) = u, 0 \leq n \leq N_{max}$. \\
		$b^*(n)$ & \makecell[l]{Optimal ordering decision with delayed orders \\ for inventory level of $n$.} \\
		$b^*_{instant}(n)$ & Same as $b^*(n)$ but for instantaneous orders. \\
		$A$ & Set of admissible pricing decision. \\
		$a_t$ & Offer price at time $t$. \\
		$a^*(n)$ & Optimal pricing decision for inventory level of $n$. \\
		$\vec{s}$ & Market situation, vector of merchant offers. \\
		%todo change to x(a, s)
		$\vec{x}(\vec{s})$ & Explanatory variables for market situation $\vec{s}$. \\
		$\lambda(\vec{s}; \vec{\beta})$ & Estimated mean sales in one period. \\
		$\vec{\beta}$ & Weights vector for linear regression. \\
		$\vec{\beta}^*$ & Optimal weights for specific trainings data. \\
		$\gamma(a)$ & \makecell[l]{Probability for a single customer of buying an item. \\ Used for Bernoulli distribution.} \\
		$V_0^{old}(n)$ & Result of value function from last computation. \\
		\bottomrule
%	\end{tabular}
	\caption{List of variables and parameters with description.}
	\label{tab:notations}
%\end{table}
\end{longtable}
