
\chapter{Merchant}
This thesis proposes a merchant software agent that can compete against competition on an online marketplaces.
The merchant makes ordering and pricing decisions with the goal to maximize the profit.
The merchant uses demand learning to estimate future customer demand from historical market situations and sales data.
The estimated demand is used in a dynamic programming approach to generate ordering and pricing policies.
%what is a policy?

We developed the merchant in four phases to overcome the complexity of this problem.
The first phase constrained the merchant's problem the most and only considers the inventory problem with known customer demand in a monopoly.
Each following phase adds a new difficulty for the merchant until the fourth phase which represents our final problem: The joint inventory and dynamic pricing problem with unknown customer behavior under competition.

Each of the following sections explains one phase in detail.
\todo{explain each ''step'' short in some sentences each with references to chapter} \cref{section:ordering}.
%explain structure of following sections

\section{Ordering}
\label{section:ordering}
In this scenario, the merchant cannot make pricing decisions.
Instead, all products are sold to a fixed price.
The merchant can focus on the ordering problem and make ordering decisions that promise the most expected long-term profit.
The merchant has a monopoly and does not have to care about any competitor.
Ordering decisions depend on the customer demand.
The demand is stochastic and its distribution is known to the merchant.
There are no backlogs. %sentence to short?
If a customer arrives when the merchant is out of stock, the merchant will miss the sale.

\subsection{Model Description}
\label{subs:ordering_model}
%situation
%todo more intro/situation
The merchant wants to sell items on the marketplace.
% price & revenue
Items are offered at a fixed price $a_{fix}$ and each sold item generates a revenue of $a_{fix}$.
% time horizon
% write about t here?
The time horizon is infinite.
% discrete time
Since in real-life applications prices cannot be adjusted arbitrarily often, we use a discrete time
model.
%discounting
We use discounting to increase the relevance of short-term profits.
A discount factor $\delta$, $0 < \delta \leq 1$, is applied to each time period.

% inventory + holding cost
%todo introduce time and periods before that
The merchant holds items in an inventory.
The random inventory level at the start of period $t$ is denoted by $N_t$.
Storing items in the inventory causes holding costs of $l$ per item per time period. % l > 0

%ordering
The merchant can reorder items to increase the inventory level.
Order decisions, which can be made once each time period, will influence the merchant's profit.
The number of items ordered at time $t$ are denoted by $b_t$, with $b_t \geq 0$.
A order of size $b_t = 0$ means that no order is made.
%todo order delay
%set depends on time, yes? no?
The set of admissible orders quantities is denoted $B_t$.
% order cost
% explain fixed/variable order cost? explain somewhere else?
Each order causes a order cost, which consists of fixed order cost $c_{fix}$ and variable order cost $c_{var}$.
%order costs paid upfront
Order costs are defined by:

$$
C(b) := \begin{cases}
c_{fix} + c_{var} \cdot b  & \quad \text{if } b > 0 \\
0  & \quad \text{if } b = 0
\end{cases}
$$

%sales
The probability to sell $i$ items is denoted by $P(i)$.
%check: myopic somewhere mentioned and explained
Note, that the probability is time-independent because of the consumer's myopic buying behavior.
Additionally, the probability in this scenario is independent from offer prices because the merchants sells in a monopoly with a fixed price.
The random number of sold items within the period $(t-1, t)$ is denoted by $X_t$.
%check: backorders mentioned in previous section
%can I express the line below with random variables?
The merchant cannot sell more items than the inventory holds, i.e. $N_t \geq X_{t+1}$. % use specific n

%whats ordering strategy?
%how to write the policy?
Depending on a given ordering policy $(b_t)_t$, the random accumulated profit from time period $t$ is:

%can I do inf sum here? or switch to avg profit?
$$
G_t := \sum_{s=t+1}^{\infty} (\delta^{s-t} \cdot (a_{fix} \cdot X_s - l \cdot N_{s-1} - C(b_{s-1}(N_{s-1})))
$$

The objective is to find a ordering policy that maximizes the expected total profit $E(G_0 | N_0)$.

% end of model descripiton, maybe write what comes next
%Sales decrease the inventory level over time and orders increase it.

\subsection{Solution Approach}
%why use dyn prog? -> optimal; why optimal?

%what? -> dyn programming
In this section, we want to derive optimal ordering policies.
% with and without delay
We use the dynamic programming approach to find the best expected profit of the stochastic control problem.
% is n defined?
The recursive value function $V_t(n)$ describes the best expected profit $E_t(G_t | N_t)$.
Since this this value function cannot be calculated over an infinite time horizon, we introduce the end time $T$ so that $t = 0, 1, 2, ..., T$.
If $T$ is sufficiently large, $V_t$ converged enough in order that the optimal ordering policy is not affected.
%V_T = 0

$N_{max}$ denotes the upper limit of the inventory level and the order decision, $0 \leq n, b \leq N_{max}$. This does not affect the optimal solution as long as $N_{max}$ is greater than the biggest inventory level that can occur in the optimal policy.

The value function with an instantaneous arrival of orders is:
%where comes i from?
% solve infinite sum over i problem: upper limit, rest probability for i >= n
\begin{equation}
\begin{split}
V_t(n) = \max_{b \in B_t} \Bigg\{
\sum_{i \geq 0} \Big(
P(i) \cdot (
a_{fix} \cdot min(i, n + b) %sales
- l \cdot (n + b) % holding cost
- C(b) % order cost
) \\
+ \delta \cdot V_{t+1}\big(min(max(n + b - i, 0), N_{max}))\big)
\Big)\Bigg\}
\end{split}
\label{eq:dyn_prog_no_delay}
\end{equation}

%check: order delay mentioned in model description?
Solving this problem with any order delay results in a huge increase in dimensions for the calculation.
The computation time would be to long to make fast and dynamic decisions on an online marketplace.
To avoid these long computation times, we assume that the order delay is exactly one period.
This solution is a heuristic for different order delays.
The value function with order delay is shown in the following equation:

\begin{equation}
\begin{split}
V_t(n) = \max_{b \in B_t} \Bigg\{
	\sum_{i \geq 0} \Big(
		P(i) \cdot (
			a_{fix} \cdot min(i, n) %sales
			- l \cdot n % holding cost
			- C(b) % order cost
		) \\
		+ \delta \cdot V_{t+1}\big(min(max(n - i, 0) + b, N_{max}))\big)
	\Big)\Bigg\}
\end{split}
\label{eq:dyn_prog}
\end{equation}

The set of order quantities $B_t$ must contain zero to allow to make no order.
The optimal ordering policy $b_t^*(n)$ is given by the arg max of \cref{eq:dyn_prog_no_delay} and \cref{eq:dyn_prog} respectively.
%address that arg max is set, do that in next sections too
The next sections will only consider the case with delayed orders.

\subsection{Evaluation}
%konvergieren der wertefunktion
%vergleiche mit startwert 0, gleicher wert, zu hoher wert
%wie schnell konvergiert es (konvergiert es)? welchen einfluss hat der startwert?
\todo{show graph of converging value function}

%told, that this is a optimal ordering policy
%deviations from that should result in worse performance (profit)
%let's compare computed optimal policy in monopoly with policy that orders one item more(less)
%hint: use bigger dif if you cannot see difference in policies
\todo{compare with merchant that buys always one more / one less}

\section{Joint Ordering and Pricing}
\label{section:joint_ordering_pricing}
%What?
In this section, the merchant is in the same situation as in \cref{section:ordering} but additionally gains control over the selling price.
With ordering and pricing decisions enabled, the merchant's actions on the marketplace are unconstrained.
%Difference in demand
The demand in this scenario depends on the chosen selling price.
A higher price usually results in less demand.
As in the previous problem, the demand distribution is known.

%Why joint:
We cannot separate the inventory and pricing problem and solve each in isolation to solve the whole problem optimally.
Ordering and pricing decisions influence each other.
Setting a price changes the demand, which requires a different ordering policy.
The other way around, an ordering policy might require certain prices for a better control over inventory levels.
This mutual influence is the reason, why ordering and pricing should be decided jointly.

\subsection{Model Description}
\label{subs:joint_model}
This model is a extension of the model from \cref{subs:ordering_model}.
%pricing decision
Instead of a fixed selling price $a_{fix}$, the merchant sets a price $a_t$ in each period $t$, with $a_t \geq 0$.
The set of admissible pricing decisions is denoted $A_t$.
%revenue
Each sold item in period $t$ generates a revenue of $a_t$.

%demand
The probability to sell $i$ items depends now on price $a$.
The merchant knows the price-dependent probability distribution.

Depending on a given pricing and ordering policy $(a_t, b_t)_t$, the random accumulated profit from time period $t$ is:

% can i reuse name from previous section?
$$
G_t := \sum_{s=t+1}^{\infty} (\delta^{s-t} \cdot (a_{s-1} \cdot X_s - l \cdot N_{s-1} - C(b_{s-1}(N_{s-1})))
$$

The objective is to find a joint pricing and ordering policy that maximizes the expected total profit $E(G_0 | N_0)$.

\subsection{Solution Approach}
\label{section:joint_solution}
As a basis, we use the dynamic programming approach shown in \cref{eq:dyn_prog}.
The Bellman equation is extended by one dimension that contains all pricing decisions $A_t$.

%todo remove t from A and B.
\begin{equation}
\begin{split}
V_t(n) = \max_{\substack{b \in B_t \\ a \in A_t}} \Bigg\{
\sum_{i \geq 0} \Big(
P(i, a) \cdot (
a \cdot min(i, n) %sales
- l \cdot n % holding cost
- C(b) % order cost
) \\
+ \delta \cdot V_{t+1}\big(min(max(n - i, 0) + b, N_{max}))\big)
\Big)\Bigg\}
\end{split}
\label{eq:dyn_prog_joint}
\end{equation}

There are three changes to the value function.
The probability is price-dependent ($P(i, a)$), the revenue in one period depends on the chosen price ($a \cdot min(i, n)$), and the maximum expected profit over all ordering decisions and pricing decisions is used.
The optimal ordering policy $b^*_t(n)$ and pricing policy $a^*_t(n)$ are given by the arg max of \cref{eq:dyn_prog_joint}.
%optimal pricing decision

\subsection{Evaluation}
\todo{show policies with increasing prices}
% anything else I can show here?

\section{Demand Learning}
\label{section:demand_learning}

This section extends the joint inventory and dynamic pricing problem from \cref{section:joint_ordering_pricing} by not providing the merchant with the demand distribution.
%the merchant, the merchant, the merchant...
The merchant needs to know the customer behavior in order to make effective ordering and pricing decisions.
The merchant can get information about the customer behavior from past customer actions.
The merchant can analyze historical market situation and sales in order to deduce future customer behavior.
%what is this dat, where is it from? (how does it look like?)
This is called demand learning.

%from data to demand distribution

%model section removed, maybe mention somewhere that problem is same, only prob is not know

\subsection{Solution Approach}
In order to overcome the inventory and dynamic pricing problem with missing demand information, we split the whole problem into two parts.
The first part is the training of a model from historical market situations and sales in order to predict sale probabilities.
These probabilities are used in the second part to create ordering and pricing policies.
%why sales prob = demand prob? check whole section
Given predicted sale probabilities, the dynamic programming solution from \cref{section:joint_solution} can be used as it is, to calculate ordering and pricing policies.

Assuming, the merchant predicts the demand correctly, the result will optimal policies.
In other cases, the quality of policies will depend on the quality of predictions. 
The separation of the problem let us focus on the prediction of the demand probability, while having a working solution for the policy creation.

%bring data into right form
In order to to use market and sales data for demand learning, it must be transformed into a usable form.
From the market data we know, at what times the market situation changed and what the market situation was.
The sales data contains information about all of a merchant's sales, including when the sales happened.
These two data sources can be combined and aggregated to a form that contains the number of sales per minute for each market situation.

%bring data into right form, part 2 feature extraction
We use regression analysis for the demand learning.
Regression analysis models and analyzes the relationship between a dependent variable and independent variables.
In this case, the dependent variable is the sales per minute metric.
Independent variables are extracted from the market situation.
In the monopoly scenario, the demand depends only on the merchant's price.
That is why for now, there is only one independent variable, the selling price.
%motivation: why use linear regression and not something else? is relationship linear?
We use linear regression to find a relationship between the independent variables $\vec{x}$ of a market situation and the sales per minute $\hat{y}$ for this situation.
The following equation shows how to predict sales per minute $y$ for a market situation.

%show how x is vector of features and constant 1 \begin{pmatrix} 1 \\ p \end{pmatrix}
\begin{equation}
\label{eq:linear_regression}
y = \vec{x}^\intercal \cdot \vec{\beta}
\end{equation}

$\vec{\beta}$ consists of the weights of the regression model.
Given the sales per minute $y_1, y_2, \ldots, y_n$ and the independent variables $\vec{x}_1, \vec{x}_2, \ldots, \vec{x}_n$ from $n$ historical market situations, the weights $\vec{\beta}$ are calculated with the least squares approach:

%todo fix on https://en.wikipedia.org/wiki/Linear_regression, ask Rainer
\begin{equation}
\vec{\beta} = \Big(\sum{\vec{x}_i \vec{x}_i^\intercal} \Big)^{-1}
			  \Big(\sum{y_i \vec{x}_i} \Big)
\end{equation}

%predicted mean demand -> distribution
The merchant is able to predict mean sales per minute for a market situation with the regression analysis.
However, sales probabilities are needed for the dynamic programming approach.
The mean sales per minute can be used as a parameter to create a probability distribution.
%find better citation with a study or something to show that.
We describe the demand distribution with a Poisson distribution.
This discrete distribution is a good approximation for the arrival and buying process of customers \cite{DBLP:journals/ior/Wolff82}.
Equation \ref{eq:probability} shows to integration of the Poisson distribution and the linear regression from \cref{eq:linear_regression} in order to create a demand probability function.

\begin{equation}
\label{eq:probability}
P(i, \vec{x}) =
	e^{\frac{-1}{\vec{x}^\intercal \vec{\beta} \cdot p}}
	(\vec{x}^\intercal \vec{\beta} \cdot p)^{-k}
	\frac{1}
	{k!}
\end{equation}
%todo fix ugly transpose

$p$ is the length of the merchant's decision period in minutes and $p > 0$.
The sales per minute are multiplied with the period length to get the number of sales in a period.
This demand probability function is used in \cref{eq:dyn_prog_joint} to calculate order and pricing policies without knowing the demand distribution.

%retraining
The merchant has access to a steadily growing collection of market situations and sales data over time.
The quality of predictions is increased by periodic trainings that utilize the new data.
There is no trainings data available at the start of a merchant's career on the marketplace.
In that case, we use a rule-based ordering strategy and an exploration pricing strategy.
The exploration pricing strategy sets random prices.
This allows the merchant to encounter many different market situations and obtain sales data for these situations.

\subsection{Evaluation}

\todo{show how good predicted percent to real percent is}


\todo{show prediction quality over number of training examples}

\section{Competition}
Until now, the merchant offered items on the marketplace in a monopoly.
This is usually not the case on real online marketplaces.
Other merchants offer the same or similar products and everyone wants to make the most attractive offer.
A common strategy to do that is to offer the cheapest price.
When multiple merchants pursue this strategy, the price will spiral downwards.
Such a commercial competition is called price war.
It is not the best strategy to always undercut the competitors because of decreasing profit margins.
Sometimes it is better to offer a higher price.
This reduces sales, but the profit margin is higher.
This section proposes a merchant that makes optimized ordering and pricing decisions under competition with the goal to maximize profit.

In the previous sections, the sale probability was only influenced by the merchant's own actions, to be specific it was influenced by the selling price that the merchant sets.
In this scenario, the merchant's sale probability is also influenced by competitor's actions.
%model description
The model is the same as in \cref{subs:joint_model} with the exception that number of sold items $X_t$ within a period depends on the merchant's offer price and the competitor's offer prices.
Merchants have access to historical market and sales data.
The sales data contains only information about a merchant's own sales, not the sales from the competitors.

%merchant must be fast
Market situations can change frequently on online marketplaces.
A merchant must able to quickly and effectively react on new market situations in order to offer an optimal price for the new situation.
Getting fast decisions is a challenge because the increasing complexity of the inventory and dynamic pricing problem over the last sections resulting in solutions with increasing computational effort.
Nevertheless, there are ways to improve the speed of the dynamic programming approach which are described in \cref{section:faster_dyn_prog}.

\subsection{Solution Approach}


%build on top of solution from previous section


%describe two possible approaches
%todo reread, and split paragraph
The inventory level was the only state on which ordering and pricing decisions were based on in the previous problems.
With competitors on the marketplace, the merchant must also consider their offers when making ordering and pricing decisions.
One possible approach is to add competitors' offers to the state in the dynamic programming calculation.
We decided against this solution because it drastically increases the state by multiple dimensions and therefore slows the computation down by multiple orders of magnitude.
Moreover, demand learning must predict transitions of competitors' offers from one time period to another.
This means that competitors' strategies must be learned, which requires a substantial amount of trainings data and a complex demand model to make sensible predictions.
Instead, we take the following approach.
We create ordering and pricing policies whenever the market situation changes for that specific new market situation.
The ordering and pricing decisions for this approach depend only on the merchant's inventory level.
This keeps the computation fast and efficient.
Additionally, only sale probabilities are predicted but not the competitors' price reactions.
This can be done by a relatively simple demand model, which can make good predictions even with few market and sales data.
The disadvantage is that new policies must be computed for each new market situation.
The next sections tackles the problem of how to quickly compute ordering and pricing policies for new market situations.

%market situation more complex -> more features
Market situations become more complex with the introduction of competition.
Previously, there was at most one active offer at a time.
Now, there can be as many offers as competing merchant.
And the number of offers can change over time when merchants are out of stock.
In \cref{section:demand_learning}, the merchant's offer price was sufficient to describe a market situation.
With competition, the sale probability of an offer depends not only on its price but also on the price of the other offers.
%todo use everywhere explanatory variables instead independent variable
New explanatory variables are necessary for demand learning to describe the more complex market situations.
We created two new explanatory variables besides the offer price in order to describe market situations with competition.
The variables are explained in detail below.

%todo bold bullet points
\begin{description}
	\item [Own price]
		This explanatory variable hold the value of the merchant's current offer price.
		It helps to explain effects on the sale probability that are based on the total price.
		For example, a higher price generally results in less demand.
	\item [Price rank]
		This variable describes the merchant's relative position in the price ranking.
		The price rank is defined by the number of competitors' offer that cost equal or less than the merchant's offer.
		Demand learning uses this to find a relationship to the sale probability.
		For example, it is probable that a merchant sells more items by having the cheapest offer instead of the second cheapest.
	\item [Price difference to cheapest offer]
		This is the difference in price from the merchant's offer to the cheapest offer.
		If the merchant has the cheapest offer, the variables value will be 0€.
		It is a relative variation of the 'own price' metric.
		This variable could explain effects like: a customer is willing to pay 5€ more than the cheapest product but not 10€ more.
\end{description}

With these explanatory variables it is possible to describe the effect of the competition and the merchant's offer price on the sale probabilities.

The predicted sale probability are used in the dynamic programming approach to compute ordering and pricing policy.
As explained above, the policies are computed for each new market situation.
%todo fix time mentioned
At this point, the computation takes around two seconds to complete.
This is to long for the merchant to quickly and frequently react on the changing market.
In the next section, we will reduce this time to allow shorter decision intervals for the merchant.

%todo dyn progamming formula with p(i,x) in demand section
%todo make x depend on price somehow

\subsection{Efficient Dynamic Programming}
\label{section:faster_dyn_prog}
In this section, we present four approaches to increase the computational efficiency of dynamic programming.
These approaches are applied to our merchant to reduce time spent on computing policies.
This allows the merchant to react faster on new market situations.
The approaches are:

\subsubsection{using good stuff}
this is a test
%adapt search space
%use last expected profit
%change number of iteration tradeoff speed performance (genauigkeit))
%quantisation (not done here)
\subsection{Evaluation}
%verlgeiche genauen Wert mit relaxierten Wert
%show price war, this is sparta!
%zeige Anpassung an verändertes Kaufverhalten?
\section{Implementation}
%extra section for adaptive dynamic programming? -> runs in 0.1 sec
%write about dummy offer?
%easy change/add of features