
\chapter{Problem Description}
\label{chapter:problem}
This thesis investigates the combined problems of dynamic pricing and inventory control on an online marketplace under competition.
%some more intro here
Merchants order durable products from a producer with unlimited supply. 
They can request the current market situation and update prices on the marketplace.
Prices can be changed over time but cannot be adjusted to different customers.
Merchants sell a single product type, but solutions for this problem are easy to extend to multiple products with independent demand.
There are no back orders. If a merchant runs out of stock, it might miss on sales.
Merchants sell products over an infinite time horizon.

%cost definition
The merchants' costs consist of fixed and variable order costs and inventory holding costs.
Fixed order costs are comparable to shipping fees and are the same per order, independent from the order size.
Variable order costs are the cost per ordered item.
Merchants pay holding costs for storing products in their inventory.
The longer an item remains in the inventory, the higher is its holding cost. 
The only source of income for a merchant are sales on the marketplace.

Inventories have no capacity limit.
However, sensible order strategies only use a limited capacity in order to save holding costs.

The objective is to find optimized ordering and pricing policies that maximize the expected profit.
A difficulty is that merchants do not know the customer behavior.
In order to adapt to the customer behavior, merchants have to observe customer actions and estimate the behavior based on these actions.
Merchants can request information about the current market situation at any time.
However, a merchant gets only data about own sales but not about competitor sales.
